SUS 2.0: Updating the System Usability Scale to conform with insights from questionnaire design research

### 2013-09-13 paper_draft1.Rnw
added more details to analysis

<<setup, include=FALSE, cache=FALSE>>=
opts_chunk$set(fig.path='plots/', fig.align='center', fig.show='asis', echo=FALSE, comment = "", message = FALSE, tidy=FALSE, dev='pdf', out.width="\\linewidth")
options(replace.assign=TRUE)
library(knitr)
library(formatR)
library(ggplot2)
library(grid)
library(boot)
library(lsr)
library(psych)
setwd("~/Dropbox/r-code/SUS2/")
m.sus2 = read.csv("sus2_medstat.csv")
m.orisus = read.csv("ori_sus_medstat.csv")
m.revsus = read.csv("rev_sus_medstat.csv")
a.sus2 = read.csv("sus2_algo.csv")
a.orisus = read.csv("ori_sus_algo.csv")
a.revsus = read.csv("rev_sus_algo.csv")
SE=function(x){
    return(sqrt(var(x)/length(x)))
}
convSUS = function(x) {
    return(rowSums((x+1)*2))
}
convSUS2 = function(x) {
    return(10*(as.matrix(x+1) %*% (c(5,5,7,7,5)/20))[,1])
}
convSUS2s = function(x) {
    return(10*(as.matrix(x+1) %*% (c(5,7,7,5)/25))[,1])
}
@

@@

\begin{abstract}
The System Usability Scale (SUS) is probably the most widely employed measure of usability. Numerous publications have been devoted to assessing its psychometric properties and using it as a gold standard in the development of alternative metrics. From a questionnaire design perspective, however, the SUS has striking deficiencies which lead to biased measures of usability. First, we review the literature on survey biases, inspect each SUS item for issues and show that the SUS is vulnerable to significant acquiescence bias using a survey experiment. We then propose the SUS 2.0, an updated version of the SUS, which conforms to insights from questionnaire design research, and present its favorable psychometric properties.
\end{abstract}

@@

\subsubsection{Internal consistency}
For the 439 responses to the original SUS, Cronbach's $\alpha$ is 0.86 and the correlation of each item with the total score lies between 0.53 and 0.81 with 95\% confidence. Although the SUS is frequently reported to have a higher coefficient alpha \cite{bangor2008,lewis2009factor,borsci2009dimensionality}, an alpha of 0.86 reflects a high degree of interrelatedness.
<<eval=FALSE>>=
alpha(m.orisus[,1:10])$total
for(i in 1:10){
    print(cor.test(rowMeans(m.orisus[,1:10]), m.orisus[,i])$conf[1:2])
}
@

\subsubsection{Factor analysis}
%Lewis \& Sauro \cite{lewis2009factor} investigated the factor structure of the SUS and found two moderately correlated factors: Usability (items 2-4, 6-10) and Learnability (items 1, 5). 
A factor analysis of our 439 responses suggest that a single-factor solution is viable, even though two factors have eigenvalues greater than one. A scree plot (Figure \ref{fig:scree}) illustrates this result.

<<eval=FALSE>>=
fa.parallel(m.orisus[,1:10])
scree(m.orisus[,1:10])
fa(m.orisus[,1:10],1)$e.values
@

@@

\begin{table}[h]
\small
\centering
\caption{Strong evidence for acquiescence bias in the SUS for individual items and the overall score}
\label{tab:acqui}
\begin{tabular}{rccccr}
\toprule
 & \multicolumn{2}{c}{Original} & \multicolumn{2}{c}{Reversed} \\
\cmidrule(r){2-3} \cmidrule(r){4-5}
\# & M & SD & M & SD & p value \\
\midrule
1 & 8.52 & 2.12 & 6.69 & 2.98 & \textless0.001\\
2 & 9.09 & 1.85 & 8.15 & 2.47 & \textless0.001\\
3 & 8.58 & 1.79 & 8.90 & 1.83 & 0.008\\
4 & 8.19 & 2.36 & 8.66 & 1.68 & \textless0.001\\
5 & 8.24 & 1.81 & 8.18 & 1.90 & 0.67\\
6 & 8.55 & 2.12 & 8.46 & 1.73 & 0.49\\
7 & 8.61 & 1.68 & 8.90 & 1.79 & 0.012\\
8 & 7.99 & 1.85 & 8.19 & 2.02 & 0.13\\
9 & 8.52 & 2.04 & 7.77 & 2.29 & \textless0.001\\
10 & 8.17 & 1.80 & 8.42 & 2.01 & 0.056\\
\midrule
 & 84.46 & 12.91 & 82.33 & 11.39 & 0.010\\
\bottomrule
\end{tabular}
\end{table}
<<eval=FALSE>>=
# testItems=function(i, oa, ra, ob, rb){
#     print(paste(i, round(t.test(oa,ra)$p.value,3), round(t.test(ob,rb)$p.value,3), sep=", "))
#     print(paste(i, paste(round(mean(oa),2), " (", round(sd(oa),2), ")", sep=""), paste(round(mean(ra),2), " (", round(sd(ra),2), ")", sep=""),
#         paste(round(mean(ob),2), " (", round(sd(ob),2), ")", sep=""), paste(round(mean(rb),2), " (", round(sd(rb),2), ")", "\\", sep=""), sep=" & "))
# }
# for(i in 1:10){
#     testItems(i, m.orisus[,i], m.revsus[,i], a.orisus[,i], a.revsus[,i])
# }
# testItems("all", rowMeans(m.orisus[,1:10]), rowMeans(m.revsus[,1:10]), rowMeans(a.orisus[,1:10]), rowMeans(a.revsus[,1:10]))

testItems=function(i, oa, ra){
    print(paste(i, round(mean(oa),2), round(sd(oa),2), round(mean(ra),2), round(sd(ra),2), round(t.test(oa,ra)$p.value,5), sep=" & "))
}
for(i in 1:10){
    testItems(i, (2*m.orisus[,i]+2), (2*m.revsus[,i]+2))
}
testItems("all", convSUS(m.orisus[,1:10]), convSUS(m.revsus[,1:10]))

@@

Distribution in Table \ref{tab:dist}

\subsubsection{Internal consistency}
Coefficient $\alpha$ is 0.77
<<eval=FALSE>>=
alpha(m.sus2[,1:5])$total
@

\subsubsection{Factor analysis}
Figure \ref{fig:scree}
<<eval=FALSE>>=
fa(m.sus2[,1:5],3)$loadings
fa(m.sus2[,1:5],2)$loadings
fa(m.sus2[,1:5],1)$loadings
@

\subsection{Sensitivity of the SUS 2.0}
[Introduce second data set]
The two online courses were offered on two distinct online platforms that shared the same core features but differed considerably in design. Based on a design heuristic evaluation of the two systems, it was determined that B had more usability problems than A. [Need to talk about how this was evaluated. E.g. \url{http://en.wikipedia.org/wiki/Heuristic_evaluation}]

Figure \ref{fig:sens} illustrates usability ratings using the original SUS and the short SUS 2.0 for two online course interfaces where one interface has an objectively higher usability level than the other. As the usability scores from both scales were not normally distributed, 95\% confidence intervals were computed from 10,000 bootstrap replicates using the adjusted bootstrap percentile method. While the original SUS is not sensitive enough to differentiate the usability of the two interfaces with 95\% confidence, the short SUS 2.0 exhibits good sensitivity. A non-parametric Wilcoxon rank sum test for each scale supports this result ($W$=18585 and $p$=0.07 for the original SUS, and $W$=74598 and $p\textless$0.001 for the short SUS 2.0)
<<eval=FALSE>>=
wilcox.test(convSUS(m.orisus[,1:10]), convSUS(a.orisus[,1:10]))
wilcox.test(convSUS(m.sus2[,2:5]), convSUS(a.sus2[,2:5]))
@

@@

<<eval=FALSE>>=
alpha(m.orisus[,1:10])
fa.diagram(fa(ori_sus,1))
fa.diagram(fa(ori_sus,2))

cor.test(oe,rowSums(ori_sus))
cor.test(difficulty,rowSums(ori_sus))
cor.test(redo,rowSums(ori_sus))
cor.test(learnsat,rowSums(ori_sus))

@@

\begin{table}
\small
\centering
\caption{Statistical information on the SUS distributions (SUS 2.0 scores scaled to 0-100 for comparison)}
\label{tab:dist}
\begin{tabular}{lccccccc}
\toprule
SUS & N & Min & Max & Mean & SD & Median & IQR\\
\midrule
Original   & 439 & 38 & 100 & 84.5 & 12.9 & 88 & 16\\
Reversed   & 438 & 26 & 100 & 82.3 & 11.4 & 84 & 14\\
Short 2.0  & 869 & 24.5 & 84.5 & 69.7 & 10.4 & 71 & 13\\ 
\bottomrule
\end{tabular}
\end{table}

<<scree, fig.width=4, fig.height=3, fig.cap='Scree plot for original SUS and SUS 2.0 showing that a single-factor solution is viable, although the original SUS has two factors with eigenvalue greater one'>>=
scree=data.frame(
    n=c(1:10,1:5), 
    eig=c(fa(m.orisus[,1:10],1)$e.values, fa(m.sus2[,1:5],1)$e.values), 
    sus=rep(c("Original SUS", "Short SUS 2.0"),c(10,5)))
ggplot(scree, aes(factor(n), eig, group=sus, linetype=sus)) + geom_point() + geom_line()  + geom_hline(yintercept=1) + theme_bw() + labs(x="Factor Number", y="Eigenvalue") + ylim(0,4.5) + theme(legend.title=element_blank(), legend.background=element_blank(), legend.key=element_blank(), legend.key.width=unit(1.5,"cm"), legend.position=c(.65,.75))
@

>>>>

<<setup, include=FALSE, cache=FALSE>>=
opts_chunk$set(fig.path='plots/', fig.align='center', fig.show='asis', echo=FALSE, comment = "", message = FALSE, tidy=FALSE, dev='pdf', out.width="\\linewidth")
options(replace.assign=TRUE)
library(knitr)
library(formatR)
library(ggplot2)
library(grid)
library(boot)
library(lsr)
library(psych)
setwd("~/Dropbox/r-code/SUS2/")
m.sus2 = read.csv("sus2_medstat.csv")
m.orisus = read.csv("ori_sus_medstat.csv")
m.revsus = read.csv("rev_sus_medstat.csv")
a.sus2 = read.csv("sus2_algo.csv")
a.orisus = read.csv("ori_sus_algo.csv")
a.revsus = read.csv("rev_sus_algo.csv")
SE = function(x){
    return(sqrt(var(x)/length(x)))
}
convSUS = function(x) {
    return(rowSums((x+1)*2))
}
convSUS2 = function(x) {
    return(10*(as.matrix(x+1) %*% (c(5,5,7,7,5)/20))[,1])
}
convSUS2s = function(x) {
    return(10*(as.matrix(x+1) %*% (c(5,7,7,5)/25))[,1])
}
@

@@

\begin{abstract}
The System Usability Scale (SUS) is probably the most widely employed measure of usability. Numerous studies have assessed its psychometric properties and used it as a `gold standard' in the development of alternative scales. From a questionnaire design perspective, however, the SUS has striking deficiencies which lead to biased measures of usability. First, we review the literature on survey biases, inspect each SUS item for issues and show that the SUS is vulnerable to significant acquiescence bias using a survey experiment. We then propose the SUS 2.0, an updated version of the SUS, which conforms to insights from questionnaire design research, and present its favorable psychometric properties. We present evidence from a second study that suggests that the SUS 2.0 is a more sensitive usability measure than the original SUS. 
\end{abstract}

@@

\begin{table}
\small
\centering
\caption{Statistical information on the SUS distributions (short SUS 2.0 scores scaled to 0-100 for comparison)}
\label{tab:dist}
\begin{tabular}{lcccccccc}
\toprule
SUS & N & Min & Max & Mean & SD & Median & IQR\\
\midrule
Original   & 439 & 38 & 100 & 84.5 & 12.9 & 88 & 16\\
Reversed   & 438 & 26 & 100 & 82.3 & 11.4 & 84 & 14\\
Short 2.0  & 869 & 24.5 & 84.5 & 69.7 & 10.4 & 71 & 13\\ 
\bottomrule
\end{tabular}
\end{table}

\subsubsection{Concurrent Validity}
We expect the original SUS to have high concurrent validity in the form of strong associations with related and weak associations with unrelated constructs. We find that the original SUS correlates weakly to moderately albeit significantly with the following related constructs: respondents' overall course experience ($r$=0.34, $t(437)$=8, $p\textless$0.001), their likelihood to take another course with the same format ($r$=0.19, $t(437)$=4, $p\textless$0.001), and their satisfaction with how much they learnt ($r$=0.27, $t(437)$=6, $p\textless$0.001). However, there we find no significant association between the SUS and the difficulty of the course ($r$=0.04, $t(437)$=0.9, p=0.4).
<<eval=FALSE>>=
cor.test(m.orisus$oe,rowSums(m.orisus[,1:10]))
cor.test(m.orisus$redo,rowSums(m.orisus[,1:10]))
cor.test(m.orisus$learnsat,rowSums(m.orisus[,1:10]))
cor.test(m.orisus$difficulty,rowSums(m.orisus[,1:10]))
@

\subsubsection{Internal consistency}
For the 439 responses to the original SUS, Cronbach's $\alpha$ is 0.86 and the correlation of each item with the total score lies between 0.53 and 0.81 with 95\% confidence. Although the SUS is frequently reported to have a higher coefficient alpha \cite{bangor2008,lewis2009factor,borsci2009dimensionality}, an alpha of 0.86 reflects a high degree of interrelatedness.
<<eval=FALSE>>=
alpha(m.orisus[,1:10])$total
omega(m.orisus[,1:10], nfactors=2)
iclust(m.orisus[,1:10],nclusters=2)
for(i in 1:10){
    print(cor.test(rowMeans(m.orisus[,1:10]), m.orisus[,i])$conf[1:2])
}
@

\subsubsection{Factor analysis}
A factor analysis of the 439 responses to the original SUS suggests that the scale has a two-factor structure. A scree plot (Figure \ref{fig:scree}) illustrates that two factors have eigenvalues greater than one. This is consistent with previous work on the SUS factor structure \cite{lewis2009factor}.
<<eval=FALSE>>=
fa.parallel(m.orisus[,1:10])
scree(m.orisus[,1:10])
fa(m.orisus[,1:10],2)$e.values
fa.diagram(fa(m.orisus[,1:10],2))
@
<<scree, fig.width=4, fig.height=3, fig.cap='Scree plot for original SUS and SUS 2.0 showing that a single-factor solution is viable, although the original SUS has two factors with eigenvalue greater one'>>=
scree=data.frame(
    n=c(1:10,1:5), 
    eig=c(fa(m.orisus[,1:10],1)$e.values, fa(m.sus2[,1:5],1)$e.values), 
    sus=rep(c("Original SUS", "Short SUS 2.0"),c(10,5)))
ggplot(scree, aes(factor(n), eig, group=sus, linetype=sus)) + geom_point() + geom_line()  + geom_hline(yintercept=1) + theme_bw() + labs(x="Factor Number", y="Eigenvalue") + ylim(0,4.5) + theme(legend.title=element_blank(), legend.background=element_blank(), legend.key=element_blank(), legend.key.width=unit(1.5,"cm"), legend.position=c(.65,.75))
@

@@

\begin{table}[h]
\small
\centering
\caption{Strong evidence for acquiescence bias in the SUS for individual items and the overall score}
\label{tab:acqui}
\begin{tabular}{rccccr}
\toprule
 & \multicolumn{2}{c}{Original} & \multicolumn{2}{c}{Reversed} \\
\cmidrule(r){2-3} \cmidrule(r){4-5}
\# & M & SD & M & SD & p value \\
\midrule
1 & 8.52 & 2.12 & 6.69 & 2.98 & \textless0.001\\
2 & 9.09 & 1.85 & 8.15 & 2.47 & \textless0.001\\
3 & 8.58 & 1.79 & 8.90 & 1.83 & 0.008\\
4 & 8.19 & 2.36 & 8.66 & 1.68 & \textless0.001\\
5 & 8.24 & 1.81 & 8.18 & 1.90 & 0.67\\
6 & 8.55 & 2.12 & 8.46 & 1.73 & 0.49\\
7 & 8.61 & 1.68 & 8.90 & 1.79 & 0.012\\
8 & 7.99 & 1.85 & 8.19 & 2.02 & 0.13\\
9 & 8.52 & 2.04 & 7.77 & 2.29 & \textless0.001\\
10 & 8.17 & 1.80 & 8.42 & 2.01 & 0.056\\
\midrule
 & 84.46 & 12.91 & 82.33 & 11.39 & 0.010\\
\bottomrule
\end{tabular}
\end{table}
<<eval=FALSE>>=
testItems=function(i, oa, ra){
    print(paste(i, round(mean(oa),2), round(sd(oa),2), round(mean(ra),2), round(sd(ra),2), round(t.test(oa,ra)$p.value,5), sep=" & "))
}
for(i in 1:10){
    testItems(i, (2*m.orisus[,i]+2), (2*m.revsus[,i]+2))
}
testItems("all", convSUS(m.orisus[,1:10]), convSUS(m.revsus[,1:10]))
@

@@

Table \ref{tab:dist} provides basic statistics on the distribution of short SUS 2.0 scores scaled to the same 0-100 score range as the original SUS to facilitate a comparison. Most notably, the SUS 2.0 scores range from about 25 to 85 for a system with good usability, while the original and reversed SUS suffer from ceiling effects which artificially constrain the distribution of scores.

\subsubsection{Concurrent Validity}
We start to establish the SUS 2.0's concurrent validity by investigating its association with related and unrelated constructs. We find that the short SUS 2.0 correlates moderately and significantly with respondents' overall course experience ($r$=0.31, $t(866)$=10, $p\textless$0.001), their likelihood to take another course with the same format ($r$=0.22, $t(867)$=7, $p\textless$0.001), and their satisfaction with how much they learnt ($r$=0.25, $t(867)$=8, $p\textless$0.001); however, the short SUS 2.0 is only marignally associated with the difficulty of the course ($r$=0.08, $t(867)$=2.5, p=0.01).
<<eval=FALSE>>=
cor.test(m.sus2$oe,rowSums(m.sus2[,1:5]))
cor.test(m.sus2$learnsat,rowSums(m.sus2[,1:5]))
cor.test(m.sus2$redo,rowSums(m.sus2[,1:5]))
cor.test(m.sus2$difficulty,rowSums(m.sus2[,1:5]))
@

\subsubsection{Internal consistency}
Cronbach's $\alpha$ for the short SUS 2.0 is 0.77 based on 869 responses and individual item correlations with the total score vary between 0.47 and 0.86 with 95\% confidence. Coefficient alpha is smaller for the short SUS 2.0, but given that coefficient alpha increases with the number of items, the ten-item SUS 2.0 will have higher reliability thant the five-item short SUS 2.0.
<<eval=FALSE>>=
alpha(m.sus2[,1:5])$total
omega(m.orisus[,1:10],nfactors=2)
for(i in 1:5){
    print(cor.test(rowMeans(m.sus2[,1:5]), m.sus2[,i])$conf[1:2])
}
@

\subsubsection{Factor analysis}
A factor analysis of 869 responses to the short SUS 2.0 suggests that a single-factor structure is most appropriate for the measure. As shown in the scree plot (Figure \ref{fig:scree}), only one factor has an eigenvalues greater than one and the slope changes considerably at the two factor point.
<<eval=FALSE>>=
fa(m.sus2[,1:5],3)$loadings
fa(m.sus2[,1:5],2)$loadings
fa(m.sus2[,1:5],1)$loadings
@

\subsection{Sensitivity of the SUS 2.0}
A good usability scale should exhibit a high level of sensitivity to reflect subtle differences in usability. We conducted a second survey study to investigate how the SUS 2.0 compares to the original SUS in terms of sensitivity. 
The two online courses were offered on two distinct online platforms that shared the same core features but differed considerably in design. Based on a design heuristic evaluation of the two systems, it was determined that B had more usability problems than A. [Need to talk about how this was evaluated. E.g. \url{http://en.wikipedia.org/wiki/Heuristic_evaluation}]

Figure \ref{fig:sens} illustrates usability ratings using the original SUS and the short SUS 2.0 for two online course interfaces where one interface has an objectively higher usability level than the other. As the usability scores from both scales were not normally distributed, 95\% confidence intervals were computed from 10,000 bootstrap replicates using the adjusted bootstrap percentile method. While the original SUS is not sensitive enough to differentiate the usability of the two interfaces with 95\% confidence, the short SUS 2.0 exhibits good sensitivity. A non-parametric Wilcoxon rank sum test for each scale supports this result ($W$=18585 and $p$=0.07 for the original SUS, and $W$=74598 and $p\textless$0.001 for the short SUS 2.0).
<<eval=FALSE>>=
wilcox.test(convSUS(m.orisus[,1:10]), convSUS(a.orisus[,1:10]))
wilcox.test(convSUS(m.sus2[,2:5]), convSUS(a.sus2[,2:5]))
@

@@

<<eval=FALSE>>=
cor.test(oe,rowSums(ori_sus))
cor.test(difficulty,rowSums(ori_sus))
cor.test(redo,rowSums(ori_sus))
cor.test(learnsat,rowSums(ori_sus))

@@
